---
title: "p8105_hw3_jd3924"
author: "Jiahe Deng"
date: "2022-10-12"
output: github_document
---

```{r}
library(readr)
library(tidyverse)
library(ggridges)
```
#problem1
```{r}
library(p8105.datasets)
data("instacart")
instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

#Problem 2
```{r}
accel_data =
  read_csv("accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(weekday_weekend= ifelse (day %in% c("Saturday","Sunday"), "weekdend", "weekday")) %>%
  select(week,day_id,day,weekday_weekend,everything()) %>%
  group_by(day_id,day) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_min",
    values_to = "amount_of_activity",
    names_prefix = "activity_")  %>%
  mutate(activity_min = as.numeric(activity_min))
accel_data
```
```{r}
accel_data %>%
  summarize(total_activity = sum(amount_of_activity)) %>%
  knitr::kable(digits = 1)
  
```

```{r}
accel_data %>%
  ggplot(aes(x = activity_min,y = amount_of_activity,color = day)) +
  geom_line()+
  labs(
    x = "Time / Day",
    y = "Amoung of Activity",
    title = "The Plot of Activity Time Courses for Each Day",
    caption = "Data from the accel_data.csv"
    )+
  theme(legend.position = "bottom")+
  scale_x_continuous(
    breaks = c(0,360,720,1080,1440),
    labels = c("0:00am", "6:00am", "12:00pm", "6:00pm", "24:00pm"),
    limits = c(0, 1440))

```
#Problem 3
```{r}
library(p8105.datasets)
data("ny_noaa")
```
```{r}
ny_noaa_df= 
  ny_noaa %>%
  janitor::clean_names() %>%
  separate(date,into = c("year","month","day"),sep = "-",convert = TRUE) %>%
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10,
    month=month.abb[.$month])
ny_noaa_df
```
```{r}
ny_noaa_df %>%
  group_by(snow) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

```{r}
ny_noaa_df %>%
  filter(month %in% c("Jan", "Jul")) %>%
  group_by(year,month,id) %>%
  summarize(avg_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = avg_tmax, color = month)) +
  geom_point() +geom_smooth(alpha = 0.5) +
  labs(
    x = "year",
    y = "temperature in degree celcius",
    title = "New York Each Stations'Average Max Temperature in January and in July",
    caption = "Data from the ny_noaa dataset")+
  theme(legend.position = "bottom")+
  scale_x_continuous(
    breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010))+
  facet_grid(. ~ month)
  
```
```{r}
ny_noaa_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_smooth(se = FALSE)+
  labs(
    title = "Maximum Temperature vs. Minimum Temperature in NY",
    x = "Minimum Temperature in degree Celcius",
    y = "Maximum Temp in degree celcius") + 
  theme(legend.position = "bottom")
```
```{r}
ny_noaa_df %>%
  filter(snow > 0, snow < 100) %>% 
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x = snow, y = year, fill = year)) +
  geom_density_ridges_gradient(alpha = 0.5) +
  labs(
    x = "snowfall in mm",
    y = "year",
    title = "The Distribution of Snowfall in New York") +
  theme(legend.position = "none")
```


