---
title: "p8105_hw3_jd3924"
author: "Jiahe Deng"
date: "2022-10-12"
output: github_document
---

```{r}
library(readr)
library(tidyverse)
library(ggridges)
```
#problem1
```{r}
library(p8105.datasets)
data("instacart")
instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

#Problem 2
```{r}
accel_data =
  read_csv("accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(weekday_weekend= ifelse (day %in% c("Saturday","Sunday"), "weekdend", "weekday")) %>%
  select(week,day_id,day,weekday_weekend,everything()) %>%
  group_by(day_id,day) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_min",
    values_to = "amount_of_activity",
    names_prefix = "activity_")  %>%
  mutate(activity_min = as.numeric(activity_min))
accel_data
```
The accelerometer data has `r nrow(accel_data)` rows and `r ncol(accel_data)` columns.
In total, this data have 50400 observations, and 6 variable.
The variable week tells the which week that the accelerometer data is being collected. 
The variable day_id tells which day that the accelerometer data is being collected.
The variable weekday_weekend tells they day is whether weekday or weekend.
The variable activity_min count the activity in minutes.
The variable amount_of_activity tells the amount of activity have been done.

```{r}
accel_data %>%
  summarize(total_activity = sum(amount_of_activity)) %>%
  knitr::kable(digits = 1)
  
```
From the table above we can see that the two lowest activity amount, which is 1440.
both happened in Saturday, and almost reach the end of the data collection.
Staying active the rest of the time. 
```{r}
accel_data %>%
  ggplot(aes(x = activity_min,y = amount_of_activity,color = day)) +
  geom_line()+
  labs(
    x = "Time / Day",
    y = "Amoung of Activity",
    title = "The Plot of Activity Time Courses for Each Day",
    caption = "Data from the accel_data.csv"
    )+
  theme(legend.position = "bottom")+
  scale_x_continuous(
    breaks = c(0,360,720,1080,1440),
    labels = c("0:00am", "6:00am", "12:00pm", "6:00pm", "24:00pm"),
    limits = c(0, 1440))

```
From the plot above, we can see that during normal bedtime, from 10 pm to 6 am, the activity level is low, which is reasonable. Also, we can see that there is always a rest period after high activity. Most of the high activity occurs in the morning and before bedtime.
In the afternoon, the activity is in the normal range.
Moreover, the most active day is on Wednesday.
#Problem 3
```{r}
library(p8105.datasets)
data("ny_noaa")
```
The ny_noaa data contains the information about New York's weather. 
It has 2595276 observations, and 7 variable. 
I separate the date into year, month,day according.
The variable id represent the station id.
According to the data, there are `r count(distinct(ny_noaa, id))` distinct stations
The variable prcp represent the precipitation
the variable snow represent the snowfall
the variable snwd represent the depth of snowfall
the variable tmax and tmin represent the maximum and minimun temperature.
```{r}
ny_noaa_df= 
  ny_noaa %>%
  janitor::clean_names() %>%
  separate(date,into = c("year","month","day"),sep = "-",convert = TRUE) %>%
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10,
    month=month.abb[.$month])
ny_noaa_df
```
prcp's unit has convert to mm
tmax and tmin's unit has convert to degree of celcius.

```{r}
ny_noaa_df %>%
  group_by(snow) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```
We can see that the most common snowfall is 0, which means that there is usually no snow in New York.
This is very close to the actual situation. New York is is not likely to have snowfall.
Even though there is a lot of missing value, there were more recorded snow-free days than NA.
In New York, snow is only possible in winter, and that is only for 3 months.

```{r}
ny_noaa_df %>%
  filter(month %in% c("Jan", "Jul")) %>%
  group_by(year,month,id) %>%
  summarize(avg_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = avg_tmax, color = month)) +
  geom_point() +geom_smooth(alpha = 0.5) +
  labs(
    x = "year",
    y = "temperature in degree celcius",
    title = "New York Each Stations'Average Max Temperature in January and in July",
    caption = "Data from the ny_noaa dataset")+
  theme(legend.position = "bottom")+
  scale_x_continuous(
    breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010))+
  facet_grid(. ~ month)
  
```
In January, most of the time, the temperature is between -10 to 10. There are few outliers.
But it does not cause a significant difference in the average maximum temperature over the years.
From 1980 to 2010, the overall trend is an upward one.
In July, most of the day, the temperature is between 20 to 35.There are few outliers too. 
For example, there is a day in July, the temperature is about 14 degree, which is not in the normal range. 
But these outliers does not affect the average minimum temperature over the years.
The temperature in July has been fairly stable, not rising and not falling trend.

```{r}
temp_max_min = 
  ny_noaa_df %>%
  drop_na() %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_smooth(se = FALSE)+
  labs(
    title = "Maximum Temperature vs. Minimum Temperature in NY",
    x = "Minimum Temperature in degree Celcius",
    y = "Maximum Temp in degree celcius") + 
  theme(legend.position = "bottom")
```
```{r}
snowfall_dist = 
  ny_noaa_df %>%
  filter(snow > 0, snow < 100) %>% 
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x = snow, y = year, fill = year)) +
  geom_density_ridges_gradient(alpha = 0.5) +
  labs(
    x = "snowfall in mm",
    y = "year",
    title = "The Distribution of Snowfall in New York") +
  theme(legend.position = "none")
```
```{r}
library(patchwork)
temp_max_min + snowfall_dist 
```

```{r}
ny_noaa_df %>%
  summarize(
    prcp_na = mean(is.na(prcp))*100,
    snow_na = mean(is.na(snow))*100,
    snwd_na = mean(is.na(snwd))*100,
    tmax_na = mean(is.na(tmax))*100,
    tmin_na = mean(is.na(tmin))*100)
```
We can see that the temperature have the most missing data.
Both tmax and tmin have about 43.71% missing data.
snwd have about 22.8% missing data.
snow have about 14.69% missing data.
prcp have the least missing data, about 5.62%.
Since there is about half data is missing for temperature.
any conclusion about temperature based on the datais not reliable. 
The conclusions about precipitation will be close to the real situation.



```

